---
title: "review_lec_1-7"
output: html_document
date: "2024-02-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R and RStudio  
R is an open source, **scripting** language. RStudio is a [GUI](https://www.computerhope.com/jargon/g/gui.htm) that is used to interact with R. You need to have R installed in order for RStudio to work.

## Look for working directory (which folder you are working in)
```{r}
getwd()
```

## Arithmetic 
### gets object
```{r}
x <- c(1,2,3,4,5)
```

### mean, median, and standard deviation
```{r}
mean(x)
median(x)
sd(x)
```

## R markdown
RMarkdown allows us to embed code in annotated chunks, show the results of analyses, and display graphical output all in one file. It is a great way of making a report which can be conveniently output to html or pdf.

## Titles and Text
Have a look at the format of this document; titles are indicated by a `#`. Larger numbers of `#'s` reduce the size of the font. Text with no `#'s` is smallest and not bold. To emphasize words, you can use _italics_ or **bold**. A line break is indicated by two spaces.  

## Links
I often add links to my RMarkdown files. For a website, include the page name in `[]` and the website address in `()`. [BIS 15L Webpage](https://jmledford3115.github.io/datascibiol/).    

Email is similar, but you need to include `mailto:` in the parentheses. Here is my email: [Joel Ledford](jmledford@ucdavis.edu).  

## 'classes of data' 
1. numeric, 2. integer, 3. character, 4. logical, 5. complex.
```{r}
my_numeric <- 42
my_integer <- 2L #adding an L automatically denotes an integer #only whole numbers
my_character <- "universe"
my_logical <- FALSE
my_complex <- 2+4i
```

clarify or specific one type of data
```{r}
my_integer <- 
  as.integer(my_numeric) #create a new object specified as an integer
```

```{r}
is.integer(my_integer) #is my_numeric an integer?
```

missing data
```{r}
is.na(my_missing)
```

```{r}
anyNA(my_missing)
```

## Data Structures
In addition to classes of data, R organizes data in different ways. Data structures include vectors, lists, matrices, data frames, and factors. 

vectors  
```{r}
numeric_vector <- c(1,2,3,4,5)
character_vector <- c("monday","tuesday","wednesday")
my_vector_sequence <- c(1:100)
```

pull out elements from one vector
```{r}
my_vector_sequence[3] # pull out the third element in this vector
my_vector_sequence<=3 #logical results (T/F)
my_vector_sequence[my_vector_sequence<=3] #pull out the elements
```

```{r}
[1,2] #pull out [row, column]
```

data matrix  
lab2_2 make data matrix example as harry potter

data frame
a data frame is similar to a spreadsheet. A data frame can store data of many different classes.  
```{r}
data.frame()
```

```{r}
names() #pull out column names
```

```{r}
dim()#dimension of data frame
```

```{r}
str() #dimension of data frame
glimpse() 
```

```{r}
data_frame$one_of_variables
```

```{r}
rbind() #row bind
cbind() #column bind
```

save as 'csv' file
```{r}
write.csv(hbirds, "hbirds_data.csv", row.names = FALSE)
```

read 'csv' from folder
```{r}
getwd()
```

```{r}
read.csv("lab4/data/file_name")
```

```{r}
class() #what is the class of data
as.factor() #from other class to factor 
levels() # group the same factors
#null means object is still a character not factor, unable to find the factors. 
```










